# Some basic R tasks {#basics}
 

By now, you should be familiar with **RStudio**'s layout. In this chapter, we'll cover basic tasks in **R**: importing data, exploring built-in datasets, summarizing data, accessing help files, and interpreting common errors.

## Using R's Built-In Datasets

Built-in datasets like `ChickWeight` can be loaded and explored:

```{r, eval = FALSE}
data(ChickWeight)  # Loads the dataset into memory
ChickWeight        # View dataset in the console
View(ChickWeight)  # Open dataset as a spreadsheet
```

To inspect structure:

```{r, eval = FALSE}
require(mosaic)
glimpse(ChickWeight)
```

This shows variable types and a preview of values. Help files are available:

```r
help(ChickWeight)
```

## .csv Files

Many datasets will be in *.csv* format. A tidy or data matrix format requires:

- Each variable in a column
- Each observation in a row
- Each value in its own cell

**Steps to save Excel as .csv**:

1. Enter data with headers in Row 1.
2. Go to *File > Save As*, choose "CSV (Comma delimited)".
3. Click "Yes" on any Excel warnings.

## Importing a .csv File

**From a URL**:

```{r data5, echo=T, eval=T, message=FALSE, collapse=T, warning=F}
mlbDataImport <- read.csv("http://www.csub.edu/~emontoya2/datasets/mlb2010.csv")
```

**From local drive**:

```{r data35, echo=T, eval=F}
ImportFromDrive <- read.csv(file.choose())
```

The dataset will appear in the Environment tab.

## Using RStudio's Import Feature

You can also go to *File > Import Dataset > From Text (base)* and choose your .csv file.

## Summarizing Data

Always load `mosaic` before analyzing:

```{r data35c, echo=T, eval=F}
require(mosaic)
```

**Basic Summary Format**: `goal(~ x, data = dataset)`

```{r data5aa, echo=T, eval=T}
glimpse(mlbDataImport)
```

**Histogram of salary**:

```{r gr2, echo=T, eval=T}
histogram(~ salary, data = mlbDataImport)
```

**Mean salary**:

```{r gras4, echo=T, eval=T}
mean(~ salary, data = mlbDataImport)
```

**Groupwise Mean**:

```{r gr4dd, echo=T, eval=T}
mean(~ salary | position, data = mlbDataImport)
```

**Boxplot by position**:

```{r gr4, echo=T, eval=T}
bwplot(~ salary | position, data = mlbDataImport)
```

## Exporting Graphs and Output

**Export a Plot**: Go to *Plots > Export*, choose your format (Image, PDF, Clipboard).

**Export Output**: Copy from Console or take a screenshot.

## Extracting Variables

Use `$` to extract variables from a dataframe:

```{r gr4a, echo=T, eval=T}
glimpse(trees)
treeVol <- trees$Volume
```

## Common Errors

**Could not find function**:
- Function not defined or misspelled
- Package not loaded

**Object not found**:
- Typo or uninitialized object

**Unmatched parentheses**:
- Missing or extra `)`

## Help Files

Use `help()` or `?` to view documentation:

```r
help(log)
?log
```

To explore functions:

```r
help.search("logarithm")
RSiteSearch("logarithm")
```

Help files explain usage, arguments, and examples. Always check them for clarification.

  
